---
title: "DATA607: Assignment 2"
output: html_document
author: Nakesha Fray
date: "2024-09-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)

```

## Introduction 

For this week's assignment, I connected to the "cunydata607sql.mysql.database.azure.com" database in SQL Workbench and created a table with 5 observations. I then connected the database to RStudio and loaded the data on to a dataframe. The table has 7 variables: name and 6 recently released movies. Each individual rated the movies they watched out of 5. The dataframe has some missing data since some individuals did not watch all 6 movies.

Part 1: Build Table
Part 2: Store data in SQL database

A code was used to generate a password for the cunydata607sql.mysql.database.azure.com database using my CUNY SPS ID. 

```{r sql password}

# Load necessary libraries
library(digest)     # For generating the hash
library(clipr)      # For copying to clipboard

# The string to hash
input_string <- '23331211'

# Generate the hash using SHA-256
hash_hex <- digest(input_string, algo = "sha256", serialize = FALSE)

# Declare a variable with only a quarter of the resulting hash
quarter_hash <- substr(hash_hex, 1, nchar(hash_hex) %/% 4)

# Save the result to the clipboard
write_clip(quarter_hash, allow_non_interactive = TRUE)

```

## Connecting Database in SQL Workbench to R

Part 3: Transfer data from SQL database to R dataframe

Connect to R using my username, password, dbname, host, and port with RMySQL. Then accessing the tables and creating a dataframe of "movie" from the tibble. 

```{r sql to r}
library(RMySQL)

mydb<- dbConnect(RMySQL::MySQL(), 
                 user='nakesha.fray11',
                 password='cb0413357786347f', 
                 dbname ='nakesha.fray11', 
                 host='cunydata607sql.mysql.database.azure.com',
                 port = 3306)

dbListTables(mydb)

movie <- tbl(mydb, "movie")%>%
  collect()

movie
```

## Missing data

Part 4: Missing data strategy

For this dataframe, we could do mean imputations based on the values recorded by other individuals in the same variable. We would be assuming that individuals would have similar ratings for the same movies. We would calculate the mean of the each variable (movie) and then substitute that value in the missing items. 

```{r - Mean Imputations}

movie2 <- movie %>% 
  clean_names() %>%
 dplyr::mutate(io2 = as.numeric(inside_out_2),
               io2 = ifelse(is.na(io2), mean(io2, na.rm = TRUE), io2))

movie2$io2

```


This is an example of mean imputation. I first changed inside_out_2 to numeric, then I used an ifelse statement to calculate the mean and replace the NA values in the variable inside_out_2 (io2) with the mean, while leaving non-missing values the same. I used this approach because the ratings for each movies appears to be somewhat similar.

## Conclusion

I was able to load a dataframe from SQL Workbench to R and then use it to do an example of mean imputations in R. I would try to create a larger, more complex table in my SQL Workbench with more variables and  different data types in the future.

